---
title: "growth_lrm"
author: "James Gately (heavily edited by Angela Larson)"
format: html
---

This code was originally written by James Gately to assist me in graphing in RStudio for the first time, and for the purpose of calculating growth rates and generation times. Majority of this was written by Gately, however it has since been edited and adapted to fit my needs. Whole code chunks exist as comments because of this, I didn't want to delete anything that was already written as I was figuring it out and could still be useful for anyone trying to look over my data. I am primarily uploading this for the use of one grad student, sorry for the messiness Maddie <3

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
lapply(c("tidyverse",
         "here",
         "janitor",
         "kableExtra",
         "cowplot",
         "ggpubr",
         "broom",
         "lubridate",
         "scales",
         "rcompanion",
         "FSA",
         "flextable",
         "magick",
         "tsibble"),
       require, character.only = TRUE)

```


```{r}
#### ------ Read-in data ------ ####

cell_growth <- read_csv(here("data", "cell_counts_MK.csv")) %>% 
  # reads in CSV file; I'm using the "here" function to avoid typing out the whole directory, but to use it, you will need to have everything within an r project
  clean_names() %>% # converts columns to lower-snake-case
  mutate(sampling_date = mdy(sampling_date)) %>% # change sampling date to a mdy format using lubridate
    mutate(date_label = case_when(sampling_date == "2024-12-05" ~ 0,
                                sampling_date == "2024-12-08" ~ 3,
                                sampling_date == "2024-12-11" ~ 6,
                                sampling_date == "2024-12-13" ~ 8),
         .after = sampling_date)  
# creates a new column, date_label, which depends on the sampling date, and, when applicable, the label column.
```


Next, we will create a timeseries plot:

```{r}
## --- Plot Cell Density Data for OAE 6 --- ##

# Cell counts graph
#count_summ_plot <- cell_growth %>%
 # select(sampling_date, label, date_label, count_cells_per_ml) %>% # selects the columns you'd like to use
 # na.omit()   # omits any "NA" values
#  group_by(sampling_date) %>% # grouped by noted column (used for summary calculations)
#  summarise(cells_per_ml_mean = signif(mean(count_cells_per_ml),3), cells_per_ml_sd = signif(sd(count_cells_per_ml), 3)) # calculates mean and sd
#the above was giving me an error regarding the select function, chatgpt gave me the below:

count_summ_plot1 <- cell_growth %>% 
  select(sampling_date, label, date_label, count_cells_per_ml) %>% 
  mutate(label = str_sub(label, end = -6)) %>% 
  na.omit() %>% 
 separate(col = label,
           into = c("label_new", "Treatment"),
           sep = "_",
           remove = FALSE)

count_summ_plot2 <- count_summ_plot1 %>% 
  select(sampling_date, date_label, label, Treatment, count_cells_per_ml) %>% 
  group_by(date_label, label, Treatment) %>% 
  summarise(count_mean = round(mean(count_cells_per_ml), 0), count_sd = round(sd(count_cells_per_ml), 0), n = n())

#count_summ_plot <- cell_growth %>%
#  dplyr::select(sampling_date, label, date_label, count_cells_per_ml) %>%
 # na.omit() %>%
 # group_by(date_label, label) %>%
  #summarise(cells_per_ml_mean = signif(mean(count_cells_per_ml),3), 
           # cells_per_ml_sd = signif(sd(count_cells_per_ml), 3))

date_label_order <- unique(count_summ_plot2$date_label)

## Create tsibble dataframe for time series

cell_growth_ts <- count_summ_plot2 %>% 
   dplyr::select(date_label, Treatment, count_sd, count_mean) %>% # selects the columns you'd like to use
   na.omit()%>%    # omits any "NA" values
  mutate(date_label = factor(date_label, levels = date_label_order))
  
   
## Make plot for cell growth
cell_count_p <- ggplot(data = cell_growth_ts, aes(x = date_label, y = count_mean, group = Treatment, color = Treatment)) +
    geom_line() +
    geom_point() +
    theme_cowplot() +
    geom_errorbar(aes(ymin = count_mean - count_sd, ymax = count_mean + count_sd), width = 0.02) +
    labs(x = "Time (Days)", y = expression(paste("Cell Density", "(cells mL"^-{1}, ")"))) +
     scale_color_manual(
        values = c("bru" = "violetred", "ctrl" = "mediumblue"),  # Assign custom colors
        labels = c("Brucite", "Control"))  # Rename legend items


cell_count_p

# Log version of previous graph
cell_count_p_log <- cell_count_p +
  scale_y_log10(breaks = 10^(3:6), labels = trans_format("log10", math_format(10^.x))) +  # log transforms your cell counts
   labs(x = "Time (days)", y = expression(paste("Log Cell Density ", "(cells mL"^-{1}, ")")))+
     annotate(geom="text", x = c(1.65, 1.65), y = c(2E5, 1.5E5), size = c(5, 5),
           label=c(expression(paste(mu[CTRL], " = 0.76 \u00B1 0.15 d"^-1, " (95% CI)")),
                   expression(paste(mu[BRU], " = 0.99 \u00B1 0.15 d"^-1, " (95% CI)")))) +
  annotate(geom="text", x = c(1.2, 1.2), y = c(9E4, 7E4), size = c(5, 5),
           label=c(expression(paste(t[CTRL], " = 0.91 \u00B1 0.08 d")),
                   expression(paste(t[BRU], " = 0.7 \u00B1 0.2 d"))))
cell_count_p_log


```


Finally, we will fit our data to a linear regression model:

```{r}
# Wrangle data
cell_counts_lm <- count_summ_plot1 %>% 
  select(sampling_date, Treatment, date_label, count_cells_per_ml) %>%
  na.omit() %>% 
  filter(!(date_label == 0))%>%  # Keep everything except mpg=21 or cyl=6
  #mutate(log_cell_count_ml = round(log(count_cells_per_ml), 2)) # log transforms cell counts
  mutate(log_cell_count_ml = log(count_cells_per_ml))



# Run lm by grouping
cell_counts_lm_summ1 <- cell_counts_lm %>% 
  #mutate(type = "ctrl", .after = date_label) %>% 
  select(date_label, Treatment, log_cell_count_ml) %>% # selects applicable columns
  group_by(Treatment) %>% # groups rows by noted column
  do(tidy(lm(log_cell_count_ml ~ date_label, data = .))) %>% # run linear regression model on dataframe  (cell counts as a function of sampling date)
  select(Treatment, term, estimate, std.error) %>% # select columns
  slice(2) %>% 
  mutate(estimate_round = round(estimate, 2),
         std.error_round = round(std.error, 2)) %>% 
  #mutate(across(estimate:std.error, funs(round(., 2)))) %>% # rounds across noted columns to the 100th place
  mutate(error_95 = 2*std.error_round,
         sp_gr_final = paste(estimate_round, " ", "(", error_95 ,")", sep="")) %>% 
  mutate(gen_time = round((log(2)/estimate), 2),
         error_gen_time = signif((log(2)/(estimate)^2)*(error_95), 1),
         gen_time_final = paste(gen_time, " ", "(", error_gen_time ,")", sep="")) %>% 
  select(sp_gr_final, gen_time_final) %>% # selects columns for table
  mutate(organism = "Emiliania huxleyi", .before = Treatment) # adds column with organism name

#print(cell_count_p_log) +
#ggplot2::annotate("Control: 0.5 ± 0.16", x = 3, y = 8, label = "Growth Rate", color = "darkred", size = 5)


```

